---
title: sathvik_thogaru_homework4
description: |
  hotel_bookings Dataset
author: sathvik_thogaru
date: 08-18-2021
output:
  distill::distill_article:
    self_contained: no
draft: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Importing data
This data set contains a single file which compares various booking information between hotels.

```{r}
library(skimr)
library(lubridate)
library(tidyverse)
```

importing the data and reading the top 5 rows 

```{r}
hotel_bookings <- read_csv("../../_data/hotel_bookings.csv")
head(hotel_bookings)
```


skim() is used to for getting summary statistics about variables in dataframe,tibbles,datatablesand vectors. It is mostly used with grouped dataframes
(source: https://cran.r-project.org/web/packages/skimr/vignettes/skimr.html)


```{r}
skim(hotel_bookings)
```

from the above summary statistics we can see there are a total of 119390 rows and 32 columns in the hotel_bookings dataset. 13 character variables, 18 numeric variables, and 1 date variable. there are a total of 4 missing values in the children variable. for the analysis now i will be using hotel, market segment, stays_in_weekend_nights and stays_in_week_nights.

## Varaible Description

hotel variable: type of hotel booked\
market segment : Market segment designation. In categories, the term “TA” means “Travel Agents” and “TO” means “Tour Operators”\
stays_in_weekend_nights : guest stayed at the hotel in weekend nights\
stays_in_week_nights : guest stayed at the hotel in week nights\
I am using the select() from the dplyr package which comes with tidyverse package and the piping for selecting columns


```{r}
(df <- hotel_bookings %>% 
  select(hotel,stays_in_weekend_nights,stays_in_week_nights,market_segment))
```

```{r}
unique(df$hotel)
unique(df$market_segment)
```

```{r}
resort_hotel <- filter(df, hotel == "Resort Hotel") 
city_hotel <- filter(df, hotel =="City Hotel") 
```


bookings in different market segments



city hotel

```{r}
(city_hotel_market_segment <- city_hotel %>% 
  count(market_segment))


(ggplot(city_hotel_market_segment, aes(market_segment,n))+
geom_histogram(fill = "indianred", stat = "identity") + 
labs(title = "City Market segment") + 
    theme(plot.title = element_text(size = 14, hjust = 0.5, face = "bold")))
```

resort hotel

```{r}
(resort_hotel_market_segment <- resort_hotel %>% 
  count(market_segment))


(ggplot(resort_hotel_market_segment, aes(market_segment,n))+
geom_histogram(fill = "indianred", stat = "identity") + 
labs(title = "Resort_Market segment") + 
theme(plot.title = element_text(size = 14, hjust = 0.5, face = "bold")))
```


```{r}
(df %>% group_by(hotel,market_segment) %>% count(market_segment))
```

```{r}
(ggplot(df, aes(hotel)) + 
  geom_histogram(fill = "indianred", stat = "count") + 
  labs(title = "Market segment") + 
    theme(plot.title = element_text(size = 14, hjust = 0.5, face = "bold"))+
  facet_wrap(vars(market_segment), scales = "free_y"))
```


how many number of days do people stay in the hotel?

Resort hotel

```{r}

resort_hotel$total_nights <- resort_hotel$stays_in_weekend_nights + resort_hotel$stays_in_week_nights
city_hotel$total_nights <- city_hotel$stays_in_weekend_nights + city_hotel$stays_in_week_nights


  
resort_number_of_days_guests_stay_data <- resort_hotel %>% 
  group_by(total_nights) %>% 
  count()

resort_number_of_days_guests_stay_data$relative_no_of_bookings <- resort_number_of_days_guests_stay_data$n/sum(resort_number_of_days_guests_stay_data$n)*100

  
city_number_of_days_guests_stay_data <- city_hotel %>% 
  group_by(total_nights) %>% 
  count()

city_number_of_days_guests_stay_data$relative_no_of_bookings <- city_number_of_days_guests_stay_data$n/sum(city_number_of_days_guests_stay_data$n)*100




```

```{r}
(plot1<-ggplot(resort_number_of_days_guests_stay_data, aes(total_nights,relative_no_of_bookings))+
   geom_bar(stat = "identity", fill="indianred"))+
  labs(x = "no of nights stayed at the hotel", y = "Guests[%]", title = "length of stay at the resort hotel")+
    theme(plot.title = element_text(size = 14, hjust = 0.5, face = "bold"))+
  xlim(c(0,30))


```




```{r}
(plot1<-ggplot(city_number_of_days_guests_stay_data, aes(total_nights,relative_no_of_bookings))+
   geom_bar(stat = "identity", fill="indianred"))+
  labs(x = "no of nights stayed at the hotel", y = "Guests[%]", title = "length of stay at the city hotel")+
    theme(plot.title = element_text(size = 14, hjust = 0.5, face = "bold"))+
  xlim(c(0,20))
```





















