---
title: "RAILROAD-COUNTY-TRACY"
description: |
  **Data breaking down US railroad employment numbers in 2012 by state and county. Data includes three variables: state, county and total employees. 2930 observations. Total number of employees ranges from 1 to over 3000. Specifically analyzing two subsets of data: large railroads (1000+ employees) and separately railroads in New England.**
#Source https://catalog.data.gov/dataset/total-railroad-employment-by-state-and-county-2012/resource/5a0b2831-23b9-4ce9-82e9-87a7d8f2c5d8 


author: Erin-Tracy
date: 2021-08-17
output:
  distill::distill_article:
    self_contained: false
draft: true
---
#*QUESTION- WHY WON'T SOURCE URL APPEAR IN DESCRIPTION?*

#**For now, Echo is TRUE, may want to go back later and change to FALSE**
```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

library("tidyverse")
library("readr")
library("ggplot2")
library("dplyr")
knitr::opts_chunk$set(fig.width = 5, fig.asp = 1/3)

```
# **Here I am reading in my CSV file.** 
#*QUESTION- WHY IS THIS FONT DIFFERENT FROM OTHER COMMENTS???*
```{r}
data<-read.csv("../../_data/railroad_2012_clean_county.csv")

#Head
head(data)
#Tail
tail(data)
#Dimensions
dim(data)
#Column Names
colnames(data)

```

#**Experimenting with Data Transformation**
#*Specifically determining average number of railroad employees with Summarise, count of railroads per state and average number of railroad employees by state.*  
```{r}
#Average of Railroad Employees for complete dataset
summarise(data,avg=mean(total_employees))

#Count of Railroads by State
count(data,state)

#Average of Employees per Railroad in each State
data %>%
  group_by(state) %>%
  summarise(avg=mean(total_employees))

```

#**Experimenting with prop.table**
```{r}
#Percentage of Railroads by State
#Confirm I understand this correctly
data%>%
  select(state)%>%
  table()%>%
  prop.table()*100
```

#**Experimenting with Filter and Arrange**
#*Specifically pulling only railroads with 1000+ employees, naming that group "Large Railroads"*
```{r}
#Filter out large railroads
filter(data,total_employees>=1000)

#Reassign large railroads
large_railroads<- filter(data,total_employees>=1000)

#Head
head(large_railroads)

#Count
count(large_railroads)

#Arrange by Total Employees 
arrange(large_railroads, desc(total_employees), state, county)

```

#**Experimenting with Select**
```{r}
#Large Railroads (1000+ employees)are located in these counties/states
select(large_railroads,"state", "county")
```
#**Experimenting with Filter, Vector and Piping**
```{r}
#Created subset of data that is just New England states, rename that group "new_england"
new_england <- filter(data, state %in% c("NH", "VT", "CT", "MA", "RI", "ME"))

#Average number of employees per railroad based on New England States only
summarise(new_england, avg=mean(total_employees))

#Count of railroads by state and then by county for New England States only
count(new_england,state)
count(new_england,county)

#Average number of employees per railroad per state
new_england %>%
  group_by(state) %>%
  summarise(avg=mean(total_employees))

```

#**Experimenting with Crosstabs**
```{r}
#This is not pretty, I'd like to revisit using ranges
xtabs(~total_employees + state,new_england)
```
#**Experimenting with ggplot, boxplot, WORK ON THIS SOME MORE! NEED LABELS**
```{r}
#Boxplot for New England States
ggplot(new_england,aes(state,total_employees))+geom_boxplot()
```

#**Experimenting with ggplot, geom_point, WORK ON THIS SOME MORE!**
```{r}
#Geompoints for New England States
ggplot(new_england,aes(state,total_employees))+
  geom_point()+
  geom_smooth()+
  labs(title = "Railroad Employee Counts by State", y = "Total Employees", x = "State") 

#Unfinished
#ggplot(data=new_england)+
  #geom_point(mapping=aes(x=county, y=total_employees))+
  
```
#**Experimenting with ggplot and fill**
#*QUESTION- CAN I MAKE KEY SMALLER AND GRAPH LARGER?*
```{r}
#Geomplot for New England States with County filled (Two Ways)
ggplot(new_england,aes(state, fill=county))+ geom_bar()+
  theme_bw()+
  labs(title="New England States Railroad Employee Counts by State and County", y="Number of Employees", x= "State")

ggplot(data=new_england)+
  geom_bar(mapping=aes(x=state, fill=county))
  theme_bw()+
  labs(title="New England States Railroad Employee Counts by State and County", y="Number of Employees", x= "State")

```

 
 
#**Experimenting with Geompoint, different dataset** 
```{r}
ggplot(data=large_railroads)+
  geom_point(mapping=aes(x=state, y=total_employees))+
  labs(title = "Railroads with 1000+ Employees", y = "Total Employees", x = "State") 

```

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.




