---
title: "Final Project"
description: |
  Baby steps
author: "Isha Akshita Mahajan"
date: 08-19-2021
output: 
  distill::distill_article:
    self_contained: false
draft: true
---

#Code Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(haven)

#load stata package and get information on variable columns 
data<- read_stata("../../_data/stata.DTA")


#load csv file and rename columns
csv<-read_csv("../../_data/abc_poll_2021.csv") %>% rename(
  "education"= "ppeduc5",
  "age"="ppage",
  "education_category"="ppeducat",
  "gender"="ppgender",                              
  "race_and_ethnicity"="ppethm",                     
  "household_size"="pphhsize",
  "household_income"= "ppinc7",
  "marital_status"="ppmarit5",
  "msa_status"="ppmsacat",
  "region_of_residence"="ppreg4",
  "living_ownership_status"="pprent",
  "state"="ppstaten",
  "work_status"="PPWORKA",
  "employement_status"= "ppemploy")
```


#Data Description


```{r}

```

#Data Questions

>1.a Do you approve or disapprove of the way Joe Biden is handling...the response to the coronavirus(COVID-19)?
>1.b Do you approve or disapprove of the way Joe Biden is handling...immigration and the situation at the United States-Mexico border?
>1.c Do you approve or disapprove of the way Joe Biden is handling...the economic recovery?
>1.d Do you approve or disapprove of the way Joe Biden is handling...the withdrawal of United States troops from Afghanistan?
>1.e Do you approve or disapprove of the way Joe Biden is handling...gun violence?
>1.f Do you approve or disapprove of the way Joe Biden is handling...crime?
>2 How concerned are you that you or someone you know will be infected with the coronavirus?...
>3 Have you personally received at least one dose of a coronavirus vaccine, or not?
>4 How would you rate the job Joe Biden is doing keeping his campaign promises?....
>5 Thinking about the next 12 months, would you say you feel optimistic or pessimistic about the way things are going in this country?

#Exploratory Analysis 


```{r}
frequnecy_analysis<- csv %>% 
  count(state)
```



