---
title: "sathvik_hotel_bookings_data_hw3"
description: |
  Hotel bookings Data
author: sathvik_thogaru
date: 08-18-2021
output:
  distill::distill_article:
    self_contained: false
draft: true
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing data
This data set contains a single file which compares various booking information between hotels.

```{r}
library(skimr)
library(lubridate)
library(tidyverse)
```

```{r}
hotel_bookings <- read_csv("../../_data/hotel_bookings.csv")
head(hotel_bookings)
```


skim() is an alternative to summary(), quickly providing a broad overview of a data frame. It handles data of all types, dispatching a different set of summary functions based on the types of columns in the data frame.

```{r}
skim(hotel_bookings)
```

## Data Wrangling
## tidying the data
finding the na values in the dataframe( row and column) using which() 

```{r}
which(is.na(hotel_bookings), arr.ind=TRUE)
```

```{r}
hotel_bookings[c(40601,40668,40680,41161),]
```

```{r}
hotel_bookings <- filter(hotel_bookings, !is.na(children)) 
```

```{r}
sum(is.na(hotel_bookings))
```

## transforming data and visualizing

```{r}
(hotel_bookings<-hotel_bookings %>% 
  mutate(arrival_month = recode(arrival_date_month,"January"=1,"February"=2,"March"=3,"April"=4,"May"=5,"June"=6,"July"=7,"August"=8,"September"=9,"October"=10,"November"=11,"December"=12)))
```


```{r}
(hotel_bookings <- hotel_bookings%>%
  mutate(arrival_date = make_date(arrival_date_year,arrival_month,arrival_date_day_of_month)))
```

looking for type of hotels booked by guests

```{r}
(hotels_info<-data.frame(table(hotel_bookings$hotel)))
```

```{r}
pie(hotels_info$Freq, labels=paste(hotels_info$Var1,sep = "=", hotels_info$Freq), main = "type of hotels booked by guests")
```


1. where do most of the guests come from?

```{r}
(guests_country_details <- hotel_bookings %>% 
   group_by(country) %>% 
   count() %>% 
   ungroup() %>% 
   arrange(desc(n)))
```

most guests by top country's

```{r}
ggplot(filter(guests_country_details, n>1500))+
  geom_bar(aes(country, n), stat = "identity")+
  labs(y="number of guests", title = "most guests by country's")+
  coord_flip()
```

2. how much do guests pay for a room at each hotel?

```{r}
city_hotel_data<-filter(hotel_bookings, hotel_bookings$hotel=="City Hotel")
resort_hotel_data<-filter(hotel_bookings, hotel_bookings$hotel=="Resort Hotel")
```

adr - average daily rate

```{r}
# city hotel
sort(unique(city_hotel_data$reserved_room_type))
# resort hotel
sort(unique(resort_hotel_data$reserved_room_type))
```

## city hotel average guests per room type 

```{r}
(city_hotel_average_guests_per_room_type <- city_hotel_data %>% 
   group_by(reserved_room_type) %>% 
   summarise(guests_per_room_type = mean(adults+children)))

```

## city hotel average daily rate per room type

```{r}
(city_hotel_adr_per_room_type_data <- city_hotel_data %>% 
   group_by(reserved_room_type) %>% 
   summarise(adr_per_room_type = mean(adr)))
```

## resort hotel average guests per room type 

```{r}
(resort_average_guests_per_room_type <- resort_hotel_data %>% 
   group_by(reserved_room_type) %>% 
   summarise(guests_per_room_type = mean(adults+children)))

```

##resort hotel average daily rate per room type

```{r}
(resort_hotel_adr_per_room_type_data <- resort_hotel_data %>% 
   group_by(reserved_room_type) %>% 
   summarise(adr_per_room_type = mean(adr)))
```

## visualizing Average daily rate of reserved room type at city hotel and resort hotel

```{r}
ggplot(city_hotel_adr_per_room_type_data, aes(x=reserved_room_type, y= adr_per_room_type))+
  geom_bar(stat = "identity",fill = "steelblue", width = .5)+
  labs(x="reserved room type", y=" average daily rate per room", title = "Average daily rate of reserved room type at city hotel")+
  theme(plot.title = element_text(size = 14, hjust = 0.5, face = "bold"))
```

```{r}
ggplot(resort_hotel_adr_per_room_type_data, aes(x=reserved_room_type, y= adr_per_room_type))+
  geom_bar(stat = "identity",fill = "steelblue", width = .5)+
  labs(x="reserved room type", y=" average daily rate per room", title = "Average daily rate of reserved room type at resort hotel")+
  theme(plot.title = element_text(size = 14, hjust = 0.5, face = "bold"))
```

 3. how does the price per night varry in the year in each hotel?

## city hotel

```{r}
city_room_prices_monthly <-select(city_hotel_data, arrival_date_month, adr)
(city_room_prices_monthly <- city_room_prices_monthly %>%
    group_by(arrival_date_month) %>%
    summarise(mean_room_prices = mean(adr)) %>% 
    ungroup() %>% 
  arrange(desc(mean_room_prices)))
```

```{r}
plot <- ggplot(city_room_prices_monthly)+
  geom_bar(aes(arrival_date_month, mean_room_prices),stat = "identity")+
  labs(x="arrival_date_month", y=" mean_room_prices", title = "mean room prices over the year at city hotel")+
  theme(plot.title = element_text(size = 14, hjust = 0.5, face = "bold"))
  

plot +
  scale_x_discrete(limits = c("January", "February", "March", "April", "May", "June", 
          "July", "August", "September", "October", "November", "December"))
```

## resort hotel

```{r}
resort_room_prices_monthly <-select(resort_hotel_data, arrival_date_month, adr)
(resort_room_prices_monthly <- resort_room_prices_monthly %>%
    group_by(arrival_date_month) %>%
    summarise(mean_room_prices = mean(adr)) %>% 
    ungroup() %>% 
  arrange(desc(mean_room_prices)))

```

```{r}
plot <- ggplot(resort_room_prices_monthly)+
  geom_bar(aes(arrival_date_month, mean_room_prices),stat = "identity")+
  labs(x="arrival_date_month", y=" mean_room_prices", title = "mean room prices over the year at resort hotel")+
  theme(plot.title = element_text(size = 14, hjust = 0.5, face = "bold"))
  

plot +
  scale_x_discrete(limits = c("January", "February", "March", "April", "May", "June", 
          "July", "August", "September", "October", "November", "December"))

```
