---
title: Saulo's Homeworks
description: |
  Feeble attempts at data science.
author:
- name: Saulo DePaula
date: 08-16-2021
categories:
- saulo
- homeworks
- read data
output:
  distill::distill_article:
    self_contained: no
draft: yes
---
# Homework Two

```{r}
eggs <- read.csv(file="../../_data/eggs_tidy.csv")
head(eggs)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework Three

### Select
```{r}
library(tidyverse)
select(eggs, "month")
```

### Filter
```{r}
library(tidyverse)
filter(eggs, `month` == "January")
```

### Arrange
```{r}
library(tidyverse)
filter(eggs, `large_half_dozen` > 130) %>%
  arrange(`large_half_dozen`)
```

### Summarize
```{r}
library(tidyverse)
summarize(eggs, mean(`large_half_dozen`))
```

# Homework Four

### Egg Data for Large Half Dozens
The following is a brief set of summary statistics from the eggs_tidy dataset, specifically looking at the price (in cents) per Large Half Dozen Eggs from January 2004 to December 2013. I have provided the mean, median, min, and max for this variable, along with a basic visualization. No data cleaning or recoding was necessary, given the provided dataset was sufficiently clean. 

```{r}
library(tidyverse)
summarize(eggs, mean(`large_half_dozen`))
```

```{r}
library(tidyverse)
summarize(eggs, median(`large_half_dozen`))
```

```{R}
library(tidyverse)
summarize(eggs, min(`large_half_dozen`))
```

```{R}
library(tidyverse)
summarize(eggs, max(`large_half_dozen`))
```

```{R}
library(tidyverse)
ggplot(eggs, aes(`large_half_dozen`)) + geom_histogram() +
  theme_minimal() +
  labs(title = "Large Half Dozen Egg Prices (in cents) | Jan. 2004 to Dec. 2013", y = "Count of Occurances", x= "Count of Eggs")
```

### Egg Data for Large Dozens
The following is a brief set of summary statistics from the eggs_tidy dataset, specifically looking at the price (in cents) per Large Dozen Eggs from January 2004 to December 2013. I have provided the mean, median, min, and max for this variable, along with a basic visualization. No data cleaning or recoding was necessary, given the provided dataset was sufficiently clean. 

```{r}
library(tidyverse)
summarize(eggs, mean(`large_dozen`))
```

```{r}
library(tidyverse)
summarize(eggs, median(`large_dozen`))
```

```{R}
library(tidyverse)
summarize(eggs, min(`large_dozen`))
```

```{R}
library(tidyverse)
summarize(eggs, max(`large_dozen`))
```

```{R}
library(tidyverse)
ggplot(eggs, aes(`large_dozen`)) + geom_histogram() +
  theme_minimal() +
  labs(title = "Large Dozen Eggs (in cents) | Jan. 2004 to Dec. 2013", y = "Count of Occurances", x= "Count of Eggs")
```

# Homework Five

```{R}
library(tidyverse)
ggplot(eggs, aes(x=`year`, y=`large_half_dozen`, col=as_factor(`extra_large_half_dozen`))) + 
  geom_point()
```

```{R}
library(tidyverse)
ggplot(eggs, aes(x=`year`, y=`large_dozen`, col=as_factor(`extra_large_dozen`))) + 
  geom_point()
```

1. *What these visualizations represent*: 
There is an equal relationship between the prices of **Large Half Dozen** and **Extra Large Half Dozen** eggs, as well as between the prices of **Large Dozen** and **Extra Large Dozen** eggs. Essentially, it appears they go up at the same rate over time; if the Large variation goes up, so too does the Extra Large variation.

2. *Why I chose this visualization approach, and what alternative approaches I considered but decided not to pursue*: 
These visualizations are very straightforward and clearly demonstrate the relationships between the Large and Extra Large egg variations. I attempted a histogram, but received a very lengthy error, which led me to feel content with the geom_point() option.

3. *What I wished, if anything, I could have executed but found limited capability to do*:
It would have been nice to produce a larger visual that included month *and* year, to spread the points out even further, but I was unsure of how to do that. I also wish there were more data to utilize, such as the price of chickens, which would be interesting to compare to the price of eggs.
