---
title: "Homework3: Functions by MM"
description: |
  A short description of the post.
Michelle Manning
date: 08-20-2021
output:
  distill::distill_article:
    self_contained: false
draft: true
---
First, I loaded in the packages and files I needed.I am using the poultry data set for this project.

```{r set up, echo=TRUE}
library(blogbuilder)
library(distill)
library(tidyverse)
poultry <- read.csv(file="poultry_tidy.csv")
library(usethis)
```

I first used the functions select, filter, and arrange. I saved sumarize for when I ran the function group_by to highlight the function's use. After those, I used rename, case_when, across, pivot_longer and wider, purrr, and lapply.

```{r, echo=TRUE}
#Functions
#select, filter, arrange
poultry2 <- poultry %>% 
  select(Product, Year, Price_Dollar) %>% 
  filter(Product == "Whole") %>% 
  arrange(Product, Price_Dollar, Year)
head(poultry2)  
head(poultry)
#Optional Functions
#group_by & summarize
poultry_grouped <- poultry2 %>% 
  group_by(Year)
summarise(poultry_grouped, Year = mean(Year))
#rename
colnames(poultry2)
poultry3 <- poultry2 %>%
  rename(Price=Price_Dollar)
#case_when  
mean(poultry3$Price)
poultry4 <- poultry2 %>%
  mutate(Above.Below.Avg = case_when(Price_Dollar >= 2.305333 ~"Above Mean", TRUE ~ "Below Mean"))
#across
poultry5 <- poultry2 %>%
  group_by(Year) %>%
  summarise(across(starts_with("Price_Dollar"), list(mean = mean, sd = sd)))
#Pivoting
poultry_wider <- poultry %>% 
  pivot_wider(names_from = "Product",values_from = "Price_Dollar")
poultry_longer <- poultry_wider %>% 
  pivot_longer(Whole:Thighs, names_to = "Product", values_to = "Price_Dollar")
#purrr
library(stringr)
avg_fun <- function(x)
purrr::map_dbl(poultry3, n_distinct)
#lapply
poultry_upper <- lapply(poultry2$Product, toupper)
str(poultry_upper)
```

I decided to use ggplot to edify my knowledge of it. Here is the poultry's Product variable's price plotted by the month and year.

```{r, echo=TRUE}
#plotting
ggplot(data = poultry, mapping = aes(x=Month, y = Price_Dollar, color = Product, group = Product))+ 
  geom_point()+ 
  geom_line() +
  facet_grid(~Year)
```