---
title: "Hanae- Homework4"
description: |
  Homework 4: Cleaning a Data set
author: "Hanae"
date: 08-21-2021
output:
  distill::distill_article:
    self_contained: false
draft: true
---
```{r include=FALSE}
library(png)
library(grid)
image<-readPNG("../../images/Hanae/Eurotourismbeforecleaning.png")
```
For this homework, I chose to work with a data set about tourism in Europe that I downloaded from Eurostat, the statistical office of the European Union which provides statistics and data on European countries. As seen below, the data set shows a monthly breakdown of nights spent at tourist accommodation establishments of European countries between January 2019 and June 2021.


```{r echo=FALSE}
grid.raster(image)
```

The original data set contains 45 sheets. The first and second sheets summarize the data. The remaining 43 sheets represent statistics about different types of touristic establishments. Each sheet contains 31 columns. The first column shows the names of the countries and the remaining 30 columns show the dates. Each cell contains the count of the nights spent by tourists. I chose to work with sheet 1 and sheet 16 only. The type of touristic establishment displayed in these  two sheets is the most general one (more specifically "Hotels; holiday and other short-stay accommodation; camping grounds, recreational vehicle parks and trailer parks"). Both sheets also have the same unit of measure (number). The difference between them is that in Sheet 1 we find information about tourists coming from a country that is foreign to the reporting country and in Sheet 16 the information shown is about tourists coming from the reporting country itself.

Note that there are some missing values in the data. Every time there is a missing value you will see ":" in the corresponding cell. The cell to its right might have the letter "c" in it which stands for "confidential". We also find "u" and "e" next to some cells with values throughout the data set. These stand for "low reliability" and "estimated" respectively. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

To clean the data, I start by loading the necessary libraries.

```{r}
library(tidyverse)
library(readxl)
library(purrr)
```

Then, as you may have noticed, the data set needs cleaning. So, I applied techniques that I have learned previously to make it more appropriate for analysis.

```{r echo = TRUE, message=FALSE} 
#I am creating a new function that will take the name of the Excel sheet as an argument. This will allow me to apply the "cleaning" to all the sheets of the excel file.
read_tourism<-function(sheet_name){
#loading the data
  read_excel("../../_data/Euro_tourism_full_dataset.xlsx", 
#I am specifying what sheet I choose to work on
             sheet = sheet_name, 
#the 11 first rows are unnecessary titles that could be skipped
             skip=11, 
#between each two date columns, the data set has an empty column. These need to be deleted.
             col_names = c("Countries",
                           "2019-01","delete","2019-02","delete","2019-03","delete",
                           "2019-04","delete","2019-05","delete","2019-06","delete",
                           "2019-07","delete","2019-08","delete","2019-09","delete",
                           "2019-10","delete","2019-11","delete","2019-12","delete",
                           "2020-01","delete","2020-02","delete","2020-03","delete",
                           "2020-04","delete","2020-05","delete","2020-06","delete",
                           "2020-07","delete","2020-08","delete","2020-09","delete",
                           "2020-10","delete","2020-11","delete","2020-12","delete",
                           "2021-01","delete","2021-02","delete","2021-03","delete",
                           "2021-04","delete","2021-05","delete","2021-06","delete")
             )%>% 
#create a new column that will contain the origin of the tourist (either foreign or domestic tourist)
    mutate("Origin"=sheet_name) %>% 
#then select only columns that are not empty
  select(!contains("delete"))%>% 
#filter out the first few rows that contain aggregated information as well as the last two rows that contain descriptive commentary.
  filter(str_detect(Countries, "Euro|Special|:", negate=TRUE))%>% 
#then filter out the last 3 columns and any empty column
  filter(Countries!="c", Countries != "e" , Countries != "u" , Countries != "") %>%
#now we want to turn columns into rows using pivot_longer.
#pivot all columns except Countries and Origin
  pivot_longer(cols = !(Countries|Origin),
#name the new column that has the months in it Dates
               names_to = "Dates",
#name the new column that had the count Nights
               values_to = "Nights")%>%
#the column Origin now has Sheet 1 and Sheet 16 which is not very user friendly. I use mutate to rename it appropriately 
  mutate(Origin = case_when(
    Origin == "Sheet 1" ~ "Foreign country",
    Origin == "Sheet 16" ~ "Reporting country"
  ))%>%
#Since my data goes from 2019 to 2021 let's just rename "Germany (until 1990 former territory of the FRG)" Germany
    mutate(recode(Countries, `Germany (until 1990 former territory of the FRG)`= "Germany"))%>%
#we can do the same thing using mutate() and case_when():
  # mutate(Countries = case_when(
  #   Countries == "Germany (until 1990 former territory of the FRG)" ~ "Germany", 
  #   TRUE ~ Countries
  # ))%>%
#and replace : by NAs
  mutate(Nights = case_when(
    Nights == ":" ~ "NA", 
    TRUE ~ Nights
  ))%>%
#note that Nights at this point is labeled as character which becomes problematic when plotting. Here I am turning it into numeric
  mutate(Nights=as.numeric(Nights))%>%
#turning Dates into date
  mutate(Dates=as.Date(paste(Dates,"-01",sep="")))
}

#at this point my function is ready. I create a new variable tourismClean that uses purrr to apply my function read_tourism to sheets 1 and 16.

tourismClean<-map_dfr(
  c("Sheet 1", "Sheet 16"),
  read_tourism)
#and this is how it looks like
tourismClean
```

Summarizing the Data

```{r echo = TRUE, message=FALSE}

#Some descriptive stats about our data set
#average nights in all countries split by tourists' origin
tourismClean%>%
  group_by(Origin)%>%
  summarize(mean(Nights, na.rm = TRUE))

#average nights in all countries split by reporting country
tourismClean%>%
  group_by(Countries)%>%
  summarize(mean(Nights, na.rm = TRUE))

#average nights in Spain, France, Italy, and Greece split by tourists' origin
tourismClean%>%
  filter(grepl('Spain|France|Italy|Greece', Countries))%>%
  group_by(Countries)%>%
  summarize(mean(Nights, na.rm = TRUE))

```

Plotting the Data

```{r echo = TRUE, message=FALSE}
#Now my data is ready. Let's graph a scatter plot of all the countries where Dates are on the x-axis and Nights on the y-axis


#scatter plot for each country separately
ggplot(tourismClean, aes(Dates, Nights))+geom_point()+facet_wrap(vars(Countries))

#graph data for Spain, France, Italy and Greece showing only tourists visiting from a foreign country
tourismClean%>%
  filter(grepl('Spain|France|Italy|Greece', Countries))%>%
  filter(Origin=="Foreign country")%>%
ggplot(aes(Dates, Nights))+labs(y="Number of nights spent in a touristic establishment")+scale_y_continuous(labels = scales::number)+
theme_bw()+geom_area()+facet_wrap(vars(Countries))

```
