---
title: "sathvik_iris_data_hw2"
description: 
  the Iris dataset
author: sathvik_thogaru
date: 08-18-2021
output:
  distill::distill_article:
    self_contained: false
draft: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

loading the iris dataset from the datasets package and running summary statistics on the iris data

```{r}
library(datasets)
summary(iris)
```

finding the column nammes of the data

```{r}
colnames(iris)
```
dim()is used to find the dimensions of the data

```{r}
dim(iris)
```

skim() is an alternative to summary(), quickly providing a broad overview of a data frame. It handles data of all types, dispatching a different set of summary functions based on the types of columns in the data frame.

```{r}
skimr::skim(iris)
```
there are 4 numeric variables and 1 factor variable which is the species. there are a total of 3 unique species in the iris

```{r}
unique(iris$Species)
```
the three species are setosa, versicolor and virginica

ploting iris 

```{r}
plot(iris)
```

