---
title: Wrangling the Australian marriage law dataset
description: |
  A short description of the post.
author: Karl Tacheron
date: '2021-08-12'
output:
  distill::distill_article:
    self_contained: no
draft: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The Australian Bureau of Statistics 2017 Marriage Law Postal survey contains data about a nationwide vote that took place by mail. The data has a few things making reading into a tibble difficult:

 - Grouped information instead of individual observations where variables appear elsewhere
 - Many extraneous & repeated calculated variables
 - Multi-index data formatted visually into blocks
 - Multi-index column names with confusing and unhelpful names
 
To make this data usable we must transform its structure in both its column layout and its rows.

We read in the Excel file's third sheet, cut it down to only the needed variables and rows, rename the columns, and remove NA values. We also remove rows that contain section totals.

```{r}
library(tidyverse)
library(readxl)

votes <- read_excel("../../_data/australian_marriage_law_postal_survey_2017_-_response_final.xls",
           sheet=3,
           skip=6)
keepcols <- colnames(votes)[c(1,2,4,11,13)]
votes <- votes[keepcols]
newcolnames <- c("Town","Yes","No","Illegible","No Response")
colnames(votes) = newcolnames 

keeprows <- 1:(dim(votes)[1]-12) # All row numbers but the last 12
votes <- votes[keeprows,]
votes <- votes %>% drop_na("Town") # Remove rows where Town is empty

# Drop junk rows containing " (Total)"
droprows <- str_detect(votes$Town,'(Total)') * c(1:dim(votes)[1])
droprows <- droprows[droprows != 0]
votes <- votes[-c(droprows),]
```

The last step is more complicated. Each observation needs a variable for is administrative "division", but this is displayed at the top of each block. These junk rows listing the parent division names must be turned into a variable for each row. 

We get the number of each row that contains `" Divisions"`.

```{r}
divrows <- str_detect(votes$Town,' Divisions') * c(1:dim(votes)[1])
divrows <- divrows[divrows != 0]
```

We also grab the name of each row and strip the end. 

```{r}
divnames <- str_replace(votes$Town[divrows]," Divisions","")
divnames
```

A nested for loop adds the Division variable for each row that follows. The loop creates a vector that contains the first and last value of each Division block in the spreadsheet, then adds the Division variable to each one in that range of rows.

```{r}
# First create an empty Division column
votes$Division <- NA

# Loop through the values in divrows, applying the Division name for each section of values
for (i in seq(1,length(divrows))) {
  divname <- divnames[i]
  start <- divrows[i]+1 # Start in the next row after the Division name
  end <- 0 # Temporary, Are values set within the if/else might be outside scope for later code?
  if (i == length(divrows)) {
    end <- dim(votes)[1] # Use the last row if at last item in divrows
  } else {
    end <- divrows[i+1]-1 # Go to right before the next value in divrows
  }
  rows_to_do <- seq(start,end)
  votes[c(rows_to_do),]$Division <- divname # Finally, print the Division to the Division variable
}
```

The rows containing the Division names are dropped. The Division column is converted into a factor, as it is categorical data.

```{r}
votes <- votes[-c(divrows),]
votes$Division <- as_factor(votes$Division)
```

We now have a table that only contains the relevant columns, and each Town has its Division as a variable.

```{r}
head(votes)
tail(votes)
```
