---
title: Cereal Data Read - Ben Lewis
description: |
  Different Cereals and their Sodium content, sugar content, and type categorized with A & C.
Your Name: Ben Lewis
date: 08-16-2021
output:
  distill::distill_article:
    self_contained: no
draft: yes
---
```{r}
library(dplyr)
library(tidyverse)
library(magrittr)
library(knitr)
```


#Homework 2

#This is me reading in data. It contains nutritional information (sugar and sodium per serving) of popular cereals. The data set can be found [here](https://artofstat.com/datasets).

```{r}
Cereal_Data <- read.csv("../../_data/cereal.csv")
```

#Homework 3

##Using the select function to see the cereal names

```{r}
brand_data <- Cereal_Data %>%
  select(`Cereal`,`Type`)
```

#Using Filter funtion to find Cereals with Sugar greater than 10

```{r}
sugar_data <- Cereal_Data
  filter(Cereal_Data,`Sugar` >= 10)

```

#Using arrange to sort by type

```{r}
type_data <- Cereal_Data
  arrange(Cereal_Data,`Type`) 
  
```

#Summarize mean sodium and sugar of cereals 

```{r}
mean_sugar_sodium <- Cereal_Data
  summarise(Cereal_Data, mean(`Sugar`), mean(`Sodium`))

```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


