---
title: Noah_Chicken_Data
description: |
  Included is a brief analysis of chicken data with a plot between year and the average price of each cut of chicken.
author:
- name: Noah Milstein
  url: http://umass.edu/sbs/dacss
date: 08-11-2021
output:
  distill::distill_article:
    self_contained: no
draft: yes
---


```{r, echo=FALSE}
library(tidyverse)
library(knitr)
library(readxl)
library(ggplot2)
```

```{r}
poultry <- read_csv("../../_data/eggs_tidy.csv")
```

```{r}
poultry 
```

Note: This code isn't running because the variables you are using aren't in the original csv that I found - but this might be because you are using the poultry file not the eggs file. Sorry I couldn't get it to work - Meredith

```{r}
#poultry %>% group_by(year, Price_Dollar, Product) %>% ggplot() +
#  geom_line(mapping=aes(y=Price_Dollar, x=Year, color=Product), na.rm=TRUE)
```

## Chicken Data Conclusion

The graph above suggests that the price of most chicken cuts remain relatively similar over time, however B/S Breast or boneless chicken breast appears to have increased in price over recent years. Thighs have also remained relatively similar

*Noah Milstein*






